---
title: "R_Based_Project"
author: "Allen Chen"
date: "April 7, 2018"
output: github_document
---

# Forecast Total number of bike trips with accessible and available data

Motivations: 
+ Knowing the demand can help structure the supply side of things
+ Forecast revenue
+ Along with the knowledge of customer and riding data, we can make more precise decisions
like the number of docking at each station, putting bikes at a station..

# Load Library
```{r warnings=FALSE}
library(dplyr)
library(ggplot2)
```

# Load Data
```{r}
# what data can we use? Let's import it!
station <- read.csv("GoBike_data/station.csv")
trip <- read.csv("GoBike_data/trip.csv")
weather <- read.csv("GoBike_data/weather.csv")
```

Let's look at the data in San Francisco during the weekends
```{r}
# how can we filter out non-SF records in trip? trip table doesn't have city column?
# hint: join trip and station table
trip$id <- trip$start_station_id
trip <- left_join(trip, station, by = c('id'))
trip <- trip %>% filter(city == 'San Francisco')
trip
```

```{r}
# Let's filter out non-SF records in weather table
unique(weather$zip_code)
weather <- weather %>% filter(zip_code == 94107)
weather
```

```{r}
# Filter out our weekdays for both trip and weather tables
trip$start_date <- as.Date(trip$start_date, '%m/%d/%Y %k')
trip$day <- as.factor(weekdays(trip$start_date))
trip <- trip %>% filter(day == 'Saturday' | day == 'Sunday')

weather$date <- as.Date(weather$date, '%m/%d/%Y')
weather$day <- as.factor(weekdays(weather$date))
weather <- weather %>% filter(day == 'Saturday' | day == 'Sunday')
```

# Get and process our Target Variable
```{r}
# Let's get total number of bike rides for every day from 2013/8/29 to 2015/8/31
num_riders <- trip %>% group_by(start_date) %>% summarise(count = n())
num_riders
```

# Get and process our features
```{r}
# Do you see any factors that might contribute to number of rides a day? Pick them out!
weather_features <- weather %>% select(date, mean_temperature_f, mean_humidity, wind_dir_degrees)
weather_features
```

```{r}
# Check the structure of the table
str(weather_features)
```

```{r}
# Any missing values across columns?
#print('date')
#sum(is.na(weather_features$date))
sapply(weather_features, function(x) sum(is.na(x)))
```

# Explore Our data and Feature Engineering
```{r}
# Let's Merge our features and target data together to create a new dataframe data
data <- merge(weather_features, num_riders, by.x = 'date', by.y = 'start_date')
# Let's drop unneccesary columns
data <- data %>% select(-c(date))
data
```

```{r}
# Let's look at the distribution of some columns
for (i in names(data)) {
  gg <- ggplot(data, aes(x = data[[i]])) +
    geom_histogram() +
    xlab(i)
  print(gg)
}
```

```{r}
for (i in names(data)) {
  gg <- ggplot(data = data, aes(x = data[[i]], y = count)) + 
    geom_point() +
    xlab(i)
  print(gg)
}

```

# Runnign Linear Regression
```{r}
# linear regression
trainingRowIndex <- sample(1:nrow(data), 0.8*nrow(data))  # Let's do a 80/20 split. This returns the indices
training_data <- data[trainingRowIndex, ]  # Get the training data
test_data <- data[-trainingRowIndex,]      # Get the testing data
model <- lm(formula = count ~ ., data = training_data)
summary(model)
```

# Residauls
```{r}
predicted_y <- predict(model, test_data)
test_data$residual <- test_data$count - predicted_y
# Let's visualize it
x <- seq(1,nrow(test_data))
ggplot(test_data, aes(x = x, y = residual)) +
  geom_point()
```

```{r}
ggplot(test_data, aes(x = residual)) +
    geom_histogram()
```

# Calculate errors
```{r}
# Implement rmse function
RMSE = function(predicted, observed){
  sqrt(mean((predicted - observed)^2))
}
RMSE(predicted_y, test_data$count)
```

```{r}
model$coefficients
```




# Just some notes
```{r}
# Feature Engineering
num_riders <- num_riders %>% mutate('day' = weekdays(start_date, abbreviate=TRUE))
ggplot(num_riders, aes(day, num_trips)) +
  geom_bar(stat='identity')


# Reset factors to be in day order
num_riders$day <- factor(num_riders$day, levels=c('Mon', 'Tue', 'Wed', 'Thu', 'Fri', 'Sat', 'Sun'))
ggplot(num_riders, aes(day, num_trips)) +
  geom_bar(stat='identity')
```

```{r}
num_riders <- num_riders %>% mutate('weekend'=(day == 'Sat' | day == 'Sun'))
ggplot(num_riders, aes(x=start_date, y=num_trips)) +
  geom_point(aes(color = weekend))
```
















